### 2023-01-11 Article.tex
Restore Correct versions of Article

% !TeX encoding = UTF-8
\documentclass[twoside,english,1p,final,sort&compress]{elsarticle}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\pagestyle{headings}
\usepackage{amsthm}
\usepackage{gensymb}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{accents}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{xcolor}
\usepackage{amssymb}
\usepackage{subcaption}
\usepackage{adjustbox}
\usepackage{placeins}
\usepackage{ulem}
\usepackage[colorlinks = true]{hyperref}

% To have Times only in text not maths
\renewcommand{\rmdefault}{ptm}

\makeatletter

\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}

% specify here the journal
\journal{International Journal of Mechanical Sciences}

% use this if you need line numbers
\usepackage{lineno}

\makeatother

\usepackage{babel}
\providecommand{\theoremname}{Theorem}

\DeclareRobustCommand{\w}{\mbox{\large\ensuremath{\mathsf{w}}}}
\DeclareRobustCommand{\dotp}{\boldsymbol{\cdot}}
\DeclareRobustCommand{\e}[1]{{\rm e}^{#1}}
\DeclareRobustCommand{\lay}[1]{^{(#1)}}
\DeclareRobustCommand{\mdot}[1]{\accentset{\mbox{\bfseries .}}{#1}}
\DeclareRobustCommand{\ie}{\emph{i.e.}\@\xspace}
\DeclareRobustCommand{\eal}{et \emph{al.}\@\xspace}
\DeclareRobustCommand{\eg}{e.g.,\@\xspace}
\DeclareRobustCommand{\RMSE}{\text{E}_\text{RMS}}
\DeclareRobustCommand{\AARE}{\text{E}_\text{AAR}}
\DeclareRobustCommand{\R}{\text{R}}
\DeclareRobustCommand{\ps}{\text{s}^{-1}}
\DeclareRobustCommand{\mr}[2]{\multirow{#1}{*}{#2}}
\DeclareRobustCommand{\MPa}{\text{MPa}}

% This is used to add comments into the text
%
% Will have to be removed from final version
%
\usepackage{soul}
\usepackage{color}
\definecolor{VWyellow}{RGB}{255,251,150}
\DeclareRobustCommand{\OP}[1]{\begingroup\sethlcolor{VWyellow}\textcolor{red}{\hl{\textbf{O.P.:} #1}}\endgroup}
\DeclareRobustCommand{\OPP}[1]{\begingroup\sethlcolor{VWyellow}\textcolor{red}{\hl{\textbf{O.P.:} In previous sentence #1}}\endgroup}

\begin{document}
\begin{frontmatter}

\title{Interpolation and extrapolation based performance measurement of Analytical and Artificial Neural Network models in predicting hot deformation behavior of a medium carbon steel}

\author[LGP]{Pierre Tize Mha}
\author[ETS]{Prashant Dhondapure}
\author[LGP]{Amèvi Tongne}
\author[ETS]{Mohammad Jahazi}
\author[LGP]{Olivier Pantalé\corref{cor1}}
\ead{Olivier.Pantale@enit.fr}
\ead[url]{http://www.enit.fr}

\cortext[cor1]{Corresponding author}

\address[LGP]{Laboratoire Génie de Production, INP/ENIT, Université de Toulouse, 47 Av d'Azereix, Tarbes, 65016, France}
\address[ETS]{École de Technologie Supérieure, 1100 Rue Notre Dame O, Montréal, QC H3C 1K3, Canada}

\begin{abstract}
\textcolor{red}{In the present work, a critical analysis of the most commonly used analytical models as well as recently introduced Artificial Neural Network based models is conducted to evaluate their prediction accuracy within and outside the experimental interval used to generate them}. To this end, the high-temperature deformation behavior of a medium carbon steel was studied over a wide range of strains ($0.0-0.7$), strain rates ($0.001~\ps-5~\ps$) and temperatures ($1050\celsius-1250\celsius$) using hot compression tests on a Geleeble $3800$ thermomechanical simulator. The experimental flow curves were then modeled using Johnson-Cook, Modified Zerilli-Armstrong, Hansel-Spittel, Arrhenius, and the recently proposed PTM model, as well as an Artificial Neural Network model with $3$ inputs, $2$ hidden layers of $15$ and $7$ neurons, respectively, and 1 output. The Mean Absolute Relative Error (MARE) and the Root Mean Square Error (RMSE) values were used to quantify the predictive accuracy of the analyzed models. The results indicated that the Johnson-Cook and Modified Zerilli-Armstrong models presented significant error, MARRE of  $21.20\%$, while the Hansel-Spittel, PTM and Arrhenius models could predict the behavior of this alloy with about $3.56\%$ error. The Artificial Neural Network model, on the other hand, showed excellent agreement, less than $0.62\%$ error, between the predicted and experimental flow curves. To further validate the performance of the above models, their ability to interpolate and extrapolate the experimental data was also tested. The Hansel-Spittel, PTM and Arrhenius analytical models showed a good interpolation and extrapolation capability with a notable superiority for the Arrhenius model compared to the other two. The Artificial Neural Networks model was however, still the most performant compared to all the models. A correlation factor is proposed to extend the applicability of some of the investigated models to deformation conditions not included in the original experimental plan.
\end{abstract}

\begin{keyword}
Johnson-Cook \sep Zerilli-Armstrong \sep Arrhenius \sep Hansel-Spittel \sep Artificial Neural Network  \sep interpolation  \sep extrapolation 
\end{keyword}

\end{frontmatter}
\linenumbers

%----------------------------------------------------------------------------------
\section{Introduction\label{sec:Introduction}}
%----------------------------------------------------------------------------------
Large size forged blocks made of medium carbon high-strength steels are extensively used in automotive industry as dies for the production of bumpers and dashboards through the plastic injection process. The manufacturing process of the large blocks starts with ingot casting, followed by open die forging, and quench $\&$ temper heat treatment process to achieve the desired mechanical properties  \cite{chadha2017deformation,chadha2018influence,murugesan2019two}. In recent years, in order to respond to the market demand, larger size forgings need to be produced. In parallel, more stringent condition on chemical homogeneity, hardness, grain size and mechanical properties from the surface to the core of the forged block are required. Among the three manufacturing steps (casting, forging, heat treatment), forging is the one where the most important microstructural changes take place and influences the final properties that could be achieved \cite{murugesan2019hybrid,chadha2020microstructure,sripada2022effect}. Open die forging process is fundamentally a hot compression process during which work strengthening effects, such as hardening (WH) and precipitation, concomitantly take place with softening phenomena such as   recovery and recrystallization under static and/or dynamic \cite{tian2022deformation,tavakoli2019ferrite}. It has also been reported that phase transformation could take place during deformation. The extent and intensity of the above strengthening or softening phenomena strongly depend on three macroscopic quantities \ie the strain $\varepsilon$, the strain rate $\mdot\varepsilon$ and the temperature $T$  \cite{ ebrahimi2017flow,Ashtiani-2012,He-2013, Changizian-2012}.

Considering the large size of the blocks, a purely experimental approach, based on trial and error, could not be used by industry and reliable predictive tools, such a Finite Element analysis (FEA) codes have been developed and are commercially available. However, the prediction reliability of such analyses is dependent on the accuracy of the material constitutive model, which describes the mutual interactions between strain, strain rate, and temperature during deformation. Precipitation and phase changes as mentioned above are considered negligible as deformation takes place in the austenite at temperatures above the dissolution temperature of most carbides. Hence most of the constitutive models depend on macroscopic parameters that influence the hardening and softening of the material. A large number of phenomenological, semi-empirical, or physical models \cite{Rusinek-2010, Shin-2010, Lin-2011, Pantale-2021} have been developed among them, the Johnson-Cook (JC) \cite{Johnson-1983}, Hansel-Spittel (HS) \cite{chadha2018approach}, and Zerilli-Armstrong (ZA) \cite{Zerilli-1987} models are the best known and most widely available in FEA codes. Despite their simplicity, each of them suffer from shortcomings: as reported by Jia \eal \cite{Jia-2021}, the JC model suffers from lack of non-coupling between strain, strain rate and temperature  while the HS model is better adapted for higher strain rate deformation conditions \cite{chadha2018approach}. To circumvent these shortcomings, several modified forms of models have been developed \cite{chadha2018approach, Rule-1998, Vural-2003, Lin-2010, Lin-2012}. However, as reported by many authors \cite{Li-2013, Zhang-2015, Zhou-2019}, even after adjusting the constants, the high temperature flow behavior, particularly when dynamic recrystallization takes place, could not be accurately predicted and none of the models is able to accurately predict flow behavior outside the experimental testing interval that was used to determine the model constants. Due to its more physics-based formulation, the ZA model and its modified form, MZA, \cite{NematNasser-2004, Lennon-2004, Muralli-2017, Cheng-2021, Muralli-2021} and the Arrhenius-type hyperbolic sine constitutive models are preferred to the JC and HS  models both for the prediction of the hot deformation behavior and microstructure analysis of the material \cite{Jonas-1969, Mostafaei-2012, Zhang-2012}. The Arrhenius formulation has been revised several times for a more accurate determination of the activation energy for high temperature deformation \cite{Slooff-2007, Lin-2008-C}. To overcome the strong dependency of the models to a specific alloy, Tize Mha \eal \cite{TizeMha-2022} recently proposed a constitutive model, PTM, whose formulation is independent of the alloy type. This formulation is based on the MZA model and polynomial functions of undefined order that are used during the identification. However, in the PTM model, the high order of the polynomial function (up to $10$ in some cases) can affect its accuracy.

Artificial Neural Network (ANN) is an approach to predict the flow stress behavior of materials without requiring a mathematical formulation of the flow law. It is therefore not necessary to postulate a mathematical expression to identify the parameters of the model. Since the behavior of materials is highly nonlinear at high temperatures, and depends on many factors that are also nonlinear, the evaluation of the flow stress behavior by an analytical model whose parameters are identified by a classical regression method is limited. Faced with these limitations, ANN models are of major interest because they are particularly suited to deal with complex and nonlinear relationships. Consequently, ANNs have been successfully applied to predict the flow stress behavior of materials under hot working conditions \cite{Lin-2008-ANN, Lu-2011-ANN, Ashtiani-2016-CSP, Stoffel-2019-NNB}. Although ANN models can predict well the material flow behavior, there is a problem with its implementation in finite element software as reported by Pantalé \eal \cite{Pantale-2021}. In fact, the implementation of constitutive model in a FEA code requires the derivatives of this model with respect to strain, strain rate and temperature.

Although progress has been made in improvement of constitutive models to better predict the flow stress behavior of materials, there is still a problem with the efficiency of all the above models. Indeed, a model is considered appropriate to the prediction of the material's behavior when its predictions and experimental results correlate well. It is questionable whether a model that correctly describes the behavior of a material in a defined experimental window can be used to accurately predict its behavior for conditions different from those it was identified for. In other words, can interpolation and extrapolation techniques be used to extent the applicability of a constitutive equation for different materials and processing windows?

The present work aims to address the above question using a combination of the recently introduced PTM model and ANN. To this end, initially, hot compression tests, simulating the open die forging process, were carried out and the generated flow curves were modeled using the above mentioned constitutive equations and an ANN model developed in this work. Then the interpolation and extrapolation capabilities of each model is evaluated and a solution is proposed that responds to the main question of the paper. The results are validated based on experimental work carried out in the present work and data obtained from the literature.

%----------------------------------------------------------------------------------
\section{Materials and Experiments\label{sec:Materials}}
%----------------------------------------------------------------------------------

%----------------------------------------------------------------------------------
\subsection{Experimental procedure}
%----------------------------------------------------------------------------------

The material used in this study is a medium carbon steel whose chemical composition is given in Table \ref{tab:Composition}.
\begin{table}[h!]
\centering
\caption{Chemical composition of medium carbon steel, Fe = balance.}
\begin{tabular}{lccccccc}
	\hline
	Element &  C   &  Mn  &  Mo  &  Si  &  Ni  &  Cr  &  Cu  \\ \hline
	Wt~\%   & 0.30 & 0.89 & 0.52 & 0.34 & 0.68 & 1.86 & 0.17 \\ \hline
\end{tabular}
\label{tab:Composition}
\end{table}
Cylindrical samples were machined with an initial diameter of $d=10$~mm and a height of $h_0=15$~mm.
Hot compression tests were performed on a Gleeble-3800 thermomechanical simulator (see Figure \ref{fig:Gleeble3800}), for the 5 temperature levels of $1050\celsius$, $1100\celsius$, $1150\celsius$, $1200\celsius$ and $1250\celsius$ with the 6 strain rates of $0.001~\ps$, $0.01~\ps$, $0.1~\ps$, $1~\ps$, $2.0~\ps$ and $5~\ps$.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\columnwidth]{Figures/Gleeble-3800}
\caption{The Gleeble-3800 thermomechanical simulator system used for this study}
\label{fig:Gleeble3800}
\end{figure}
Thin tantalum sheets were used as a lubricating material at the contact surface of the anvils and samples to minimize friction during testing. As shown in Figure \ref{fig:GleebleProcess}, the samples were heated to a temperature of $1260\celsius$ with a heating rate of $2$\celsius/s and held at this temperature for $5$~min to eliminate thermal gradients. They were then cooled down with a rate of $1$\celsius/s to the test temperature and then held at constant temperature for $1$~min before deformation. During the compression phase, the temperature of the specimen is kept constant by the thermal control system of the machine. After compression, the specimen is quickly quenched to freeze its microstructure for later analysis. Figure \ref{fig:GleebleProcess} also shows the aspects of the specimens before and after the compression test: $h_0$ and $r_0$ are the height and radius before compression and $h$, $r_m$ and $r_t$ are the height, large radius and small radius of the specimen after compression, respectively.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.8\columnwidth]{Figures/GleebleProcess}
\caption{Schematic diagram of the experimental process.}
\label{fig:GleebleProcess}
\end{figure}

The stress-strain curves are automatically exported from the Gleeble thermomechanical simulator system as true stress and true strain according to L-gauge where the formula to get those curves is given by $\sigma=F/A$ and $\varepsilon=\ln\left(h_0+\Delta h / h_0\right)$ or C-gauge whose formula are $\sigma = 4F/\pi(d+\Delta d)^2$ and $\varepsilon = 2\ln(d/(d+\Delta d))$ with $d = 2r_0$ and $F$ is the force as measured by Gleeble load cell. As there are noises from raw data, savgolfilter method from scipy library is used to remove noise and have more smooth data. For further usage of data in numerical simulation elastic parts have been removed.

%----------------------------------------------------------------------------------
\subsection{Compression tests results\label{sec:ComTestResults}}
%----------------------------------------------------------------------------------

The set of flow stress $\sigma^y$ versus strain $\varepsilon$ curves obtained from compression tests performed on the Gleeble-3800 simulator for each test condition (6 strain rates and 5 temperatures) is presented in Figure \ref{fig:RawData}.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]{Figures/rawData}
\caption{Stress-strain curves of medium carbon alloy extracted from the Gleeble device at various temperatures $T$ and strain rates $\mdot\varepsilon$.}
\label{fig:RawData}
\end{figure}
All data curve contains $700$ equidistant strain values up-to $\varepsilon=0.7$.
So, there are $6$ strain rates and $5$ temperatures, the database consists of $21~000$ data points.
The overall behavior of these curves shows that the flow stress $\sigma^y$ increases with increasing strain rate $\mdot\varepsilon$ but decreases with increasing temperature $T$.
It should be noted that the strain also influences the flow stress.
Indeed, for the lowest strain rates $\mdot\varepsilon$, the flow stress $\sigma^y$ increases with the strain $\varepsilon$ until a value of about $\varepsilon=0.2$ to $0.3$, then decreases to maintain a more or less constant value until the end of the test.
For the highest strain rates ($1~\ps$, $2~\ps$, and $5~\ps$), the flow stress increases throughout the test.
The slight increase in stress at low strain rates, when the strain is large, has been reported to be due to friction between the sample and the anvil during the test \cite{buckley2001deformation}. The frictional effect is also visible when testing at low strain rates, as the effect of lubrication decreases over time and thus friction increases. The increase of stresses observed at the beginning of the deformation up to $0.1$ is due to the work hardening (WH). After $0.1$ up to $0.2$ the flow stress shows a continuous reduction with increasing stress until a peak or an inflection of work-hardening rate. This behavior describes that thermal softening becomes more and more predominant until it exceeds WH. At this step, the stress curve shows three different patterns with the increasing strain: (i) gradual decrease to a steady state with DRV/DRX softening. This is the case for all deformation temperatures and strain rates between $0.001$ and $0.1$ \text{s}$^{-1}$ except those at $1050^\circ$C and $1100^\circ$C; (ii) higher stress levels without significant softening and work-hardening at $1050^\circ$C and $1100^\circ$C and strain rate of $0.1$ $\text{s}^{-1}$ ; and (iii) continuous increase with significant work hardening (all deformation temperatures and strain rate of $1$ \text{s}$^{-1}$). Therefore, it can be concluded that the softening due to DRX, characterized by a flow curve with a single peak followed by a steady-state flow, takes place at high temperatures and low strain rates. In contrast, at higher strain rates and lower temperatures, the higher work hardening rate slows down the rate of softening due to DRX, and therefore, both the peak stress and the onset of steady-state flow are shifted to higher strain levels. In fact, the drop observed in
stress is because of DRX occurrence at all temperatures and strain rates of $0.001$–$5$ $\text{s}^{-1}$. As an example, Figure \ref{fig:Micrography} shows the microstructures of medium carbon held during $5$~min at $1260\celsius$ before and after the compression phase.
\begin{figure}[!ht]
\centering
\begin{subfigure}[b]{0.45\columnwidth}
\centering
\includegraphics[width=\columnwidth]{Figures/BeforeCompM}
\caption{Before hot compression}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.45\columnwidth}
\centering
\includegraphics[width=\columnwidth]{Figures/AfterCompM}
\caption{After hot compression}
\end{subfigure}
\caption{Optical micrographs of a medium carbon steel before (a) and after (b) hot compression.}
\label{fig:Micrography}
\end{figure}
As observed on the flow curves reported in Figure \ref{fig:RawData}, it can be seen by comparing the two micrographs in Figure \ref{fig:Micrography} that recrystallization is almost complete for this temperature and strain rate within the sample held for 5 minutes at $1260^\circ$C and rapidly cooled to get autenized grain in one hand and in other hand held for $1$ minute at $1150^\circ$C before compression with strain rate $0.001~\text{s}^{-1}$.
From the results reported in Figure \ref{fig:RawData}, the identification of the parameters of the 6 proposed flow laws will be presented in the next section in order to predict the behavior of the material in any test condition.

%----------------------------------------------------------------------------------
\section{Identification of constitutive flow laws parameters\label{sec:ConstLaws}}
%----------------------------------------------------------------------------------
%----------------------------------------------------------------------------------
\subsection{The Johnson-Cook model\label{sec:JC}}
%----------------------------------------------------------------------------------

The JC model, as mentioned above, is one of the most widely used analytical models because it can be applied to many materials under different conditions of strain $\varepsilon$, strain rate $\mdot\varepsilon$, and temperature $T$.
However, the formulation of this model does not take into account the simultaneous effect of strain, strain rate and temperature.
Indeed, it is formulated by describing the effect of each physical parameter ($\varepsilon$, $\mdot\varepsilon$ and $T$) separately as a factor in the mathematical expression of the model.
Hence its inability to describe the phenomenon of softening induced by temperature.
The equation that describes this model is given as follows \cite{Johnson-1983}:
\begin{equation}
\sigma^y=\left(A+B\varepsilon^{p^{n}}\right) \left[1+C\ln\left(\frac{\mdot\varepsilon}{\mdot\varepsilon_0}\right)\right] \left[1-\left(\frac{T-T_0}{T_m-T_0}\right)^{m}\right],\label{eq:Johnson-Cook}
\end{equation}
where $\sigma^y$ is the flow stress, $\varepsilon^p$ is the plastic strain, $A$ is the initial elastic limit of the material, $B$ is the strain hardening coefficient, $n$ is the strain hardening exponent, $C$ and $m$ are the material constants that describe the strain rate hardening coefficient and the thermal softening coefficient, respectively.
The other values are reference ones, so that $\mdot\varepsilon_0$ is the reference strain rate, $T_m$ and $T_0$ are the melting temperature ($1460\celsius$ in our case) and the reference temperature, respectively.
For the determination of the parameters of the analytical models, the reference values for strain rate and temperature are $\mdot\varepsilon_0=0.001~\ps$ and $T_0=1050\celsius$ respectively.

The procedure used for the determination of the parameters of the Johnson-Cook law is in accordance with the one proposed by He \eal \cite{He-2013}.
This method allows obtaining one after the other the 5 parameters in the order $A$, $B$, $n$, $C$ and $m$.
Thus, according to our experimental data, the initial elastic limit of the material at the reference strain rate $\mdot\varepsilon_0$ and the reference temperature $T_0$ is $A=13.5143~\MPa$.
For the determination of the constants $B$ and $n$, from the results of the compression test at $T_0$ and $\mdot\varepsilon_0$, it is then possible to determine these two constants by considering only the first term $\left(A+B\varepsilon^{p^{n}}\right)$ in Equation (\ref{eq:Johnson-Cook}).
Thus, here, $B=21.816~\MPa$ and $n=0.0746$.
Once the parameters $A$, $B$ and $n$ are known, the determination of $C$, considering all the curves at $T=T_0$ then gives $C=0.3404$.
Finally, the last parameter $m$ is identified from the curves at $\mdot\varepsilon=\mdot\varepsilon_0$ and from the knowledge of the previous and gives $m=0.7057$.
All parameters of the Johnson-Cook model are reported in Table \ref{tab:JC}.

\begin{table}[h!]
\centering
\caption{Parameters' values of the Johnson-Cook flow law for the P20 steel.}
\begin{tabular}{ccccc}
	\hline
	$A~(\MPa)$ & $B~(\MPa)$ &   $n$    &   $C$    &   $m$    \\
	   $13.5143$     &     $21.816$     & $0.0746$ & $0.3404$ & $0.7057$ \\ \hline
\end{tabular}
\label{tab:JC}
\end{table}

The values predicted by the Johnson-Cook flow law (solid line) and the experimental values (dots) are compared in Figure \ref{fig:CompExp-JC-6}.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]{Figures/CompExp-JC-6}
\caption{Comparison between the experimental and predicted flow stresses by JC model.}
\label{fig:CompExp-JC-6}
\end{figure}
The JC model is not able to describe the evolution of the flow stress of medium carbon for all strain values, and the predicted stresses $\sigma^y$ show a monotonically decreasing trend for several values of the strain rates.
The discrepancy between the predicted and experimental values is large for low strains and sometimes acceptable for high strains.
As expected, the mathematical formulation of the Johnson-Cook flow law is unable to represent the stress drop at low strain rates, the JC model being only monotonically increasing with strain.
Since most of the parameters are identified at low strain rates ($A$, $B$ and $n$), this results in a very poor fit of the model to the experimental data.

The accuracy and predictive ability of the models are usually assessed by certain coefficients such as the Mean Absolute Relative Error ($\AARE$) defined by Equation (\ref{eq:AARE}):
\begin{equation}
\AARE(\%) = \frac{1}{N} \sum_{i=1}^{N}{\left|\frac{\sigma_i^p -\sigma_i^e}{\sigma_i^e}\right|} \times 100, \label{eq:AARE}
\end{equation}
and the Root Mean Square Error ($\RMSE$) defined by Equation (\ref{eq:RMSE}):
\begin{equation}
\RMSE (\MPa) = \sqrt{\frac{1}{N} \sum_{i=1}^{N} \left(\sigma_i^p - \sigma_i^e\right)^2}, \label{eq:RMSE}
\end{equation}
where $\sigma_i^e$ is the experimental value and $\sigma_i^p$ is the predicted value using the given model of the flow stress $\sigma^y$ and $N$ is the total number of data points used to compute those coefficients (in our case $N=21~000$).
For the JC model, $\AARE=14.05~\%$ and $\RMSE=12.00~\MPa$.
As reported by Phaniraj \cite{Phaniraj-2003}, the correlation coefficient ($\R$) is generally not a good measure in our case of study because it only shows the correlation of the model with respect to the data and not its accuracy, which is a determining factor in the qualification of a model.
Therefore, as a good (high) value of $\R$ (close to $1$) does not necessarily mean a good prediction of the model but simply establishes a good linearity correlation between the experiment and the prediction we avoided its use in our analysis.

%----------------------------------------------------------------------------------
\subsection{The Modified Zerilli-Armstrong model\label{sec:MZA}}
%----------------------------------------------------------------------------------

The MZA which is the modified form of ZA model, as the JC model is one of widely used models and implemented in many FEM codes like Abaqus.
The difference between the MZA model and JC model is related to the consideration of the three physical parameters to describe the reality observed from experimental.
Indeed, in the JC model, those parameters are considered separately where they are simultaneously considered in MZA and for that formulation MZA is preferred to JC model \cite{Hull-2011}.
But the original form of ZA model has some limitations due to the fact that it is considered as two terms and to ameliorate that formulation Samantaray \eal \cite{Samantaray-2009} proposed a modified form given by the following equation:
\begin{equation}
\label{eq:MZA-model}
\sigma^y = \left(C_{1}+C_{2}\varepsilon^{p^n}\right) \exp\left[-\left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right) + \left(C_{5}+C_{6}\left(T-T_0\right)\right)\ln\left( \frac{\mdot\varepsilon}{\mdot{\varepsilon}_0}\right)\right],
\end{equation}
where the 7 coefficients $C_i$ and $n$ are the parameters of the model to be identified for a given material.
To get the parameters of the MZA model we apply the medod proposed by \cite{Samantaray-2009} and those parameters are summurizedin in Table \ref{tab:MZA} while its prediction are plotted in Figure \ref{fig:CompExp-MZA-6}.
\begin{table}[h!]
\centering{}
\caption{Parameters' constants of Modified-Zerilli-Armstrong Model}
\begin{tabular}{ccccccc}
	\hline
	  $C_1$   &   $C_2$   &         $C_3$          &         $C_4$          &  $C_5$   &         $C_6$         &   $n$    \\
	$13.5143$ & $21.2591$ & $4.7902\times 10^{-3}$ & $1.4895\times 10^{-4}$ & $0.1389$ & $1.495\times 10^{-4}$ & $0.0621$ \\ \hline
\end{tabular}
\label{tab:MZA}
\end{table}
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-MZA-6}
\caption{Comparison between the experimental and predicted flow stresses by MZA model}
\label{fig:CompExp-MZA-6}
\end{figure}
It can be seen from those figures that the MZA model as JC model is not suitable to describe experimental at low strain rates but much better for high strain rates.
The deviation between the predicted values and the experimental values is large because this model has problem to correctly describe the softening in its formulation.
For the MZA model, $\AARE=21.20~\%$ and $\RMSE=19.57~\MPa$, this shows an overall worse performance than the JC model as reported in \cite{TizeMha-2022}.

%----------------------------------------------------------------------------------
\subsection{The Hansel-Spittel model\label{sec:HSmodel}}
%----------------------------------------------------------------------------------

The Hansel-Spittel model \cite{Hensel-1978} is one of the least known models in terms of integration in FEM codes, although the determination of its parameters is easier than that of the JC or MZA model.
The programming of a simple identification script is sufficient to identify its parameters for a given material.
The equation of the HS model is given by the following relation:
\begin{equation}
\sigma^y = A~\e{m_1T}~\varepsilon^{m_2}~\mdot\varepsilon^{m_3}~\e{({m_4}/{\varepsilon})}~(1+\varepsilon)^{m_5T}~\e{m_6\varepsilon}~\mdot\varepsilon^{m_7T}~T^{m_8}, \label{eq:HS}
\end{equation}
where again, $\sigma^y$ is the flow stress, $\varepsilon$ is the strain, $\mdot\varepsilon$ is the strain rate and $T$ is the temperature as proposed earlier.
The coefficients $A$ and $m_i$ are the 9 parameters of the model to be identified.
However, this model has some shortcomings, notably related to the fact that its accuracy varies according to the number of parameters taken into account during the identification.
For its identification, several authors restrict its expression to a reduced number of 5 or 6 $m_i$ parameters only, by forcing a zero value for the other parameters \cite{Chadha-2018, Mehtedi-2015, Rudnytskyj-2020}.

In the present study, the best results were obtained by taking the model defined by only the first 7 $m_i$ terms of the equation (\ref{eq:HS}), so that $m_8=0$.
From the experimental data obtained during the compression tests, an identification procedure based on the use of the LMFIT optimizer \cite{Newville-2016} has allowed to compute the parameters reported in Table \ref{tab:HS}.

\begin{table}[h!]
\centering
\caption{Parameters' values of the Hansel-Spittel flow law for the P20 steel.}
\begin{tabular}{ccccc}
	\hline
	        $A$          &         $m_1$          &   $m_2$   &        $m_3$         &         $m_4$         \\
	$5.954\times 10^{3}$ & $-3.3576\times10^{-3}$ & $0.2641$  &      $-0.0868$       & $2.2688\times10^{-4}$ \\ \hline
	                     &         $m_5$          &   $m_6$   &        $m_7$         &         $m_8$         \\
	                     & $-4.2163\times10^{-4}$ & $-0.0561$ & $2.264\times10^{-4}$ &          $0$          \\ \hline
\end{tabular}
\label{tab:HS}
\end{table}

The comparison of the values predicted by the HS model and the experimental values is presented in Figure \ref{fig:CompExp-HS-6} where the dots represent the experimental values and the solid lines are the values predicted by the Hansel-Spittel flow law.
For the HS model, $\AARE=7.75~\%$ and $\RMSE=3.80~\MPa$.
It appears that this model as well as the previous ones do not predict well the experimental one and the difference is relatively important for all strain rates below $1~\ps$.
This shows that this model is not appropriate for the characterization of this alloy, in particular because of the strong non-linearity of the behavior for low values of the strain rate.
The DRX phenomena cannot be reproduced by this type of model.

\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-HS-6}
\caption{Comparison between the experimental and predicted flow stresses by HS model}
\label{fig:CompExp-HS-6}
\end{figure}

%----------------------------------------------------------------------------------
\subsection{The Arrhenius model\label{sec:ARmodel}}
%----------------------------------------------------------------------------------

The Arrhenius type model \cite{Sellars-1966} is one of the most used models in the framework of material forming, especially when it comes to studying the microstructure of the material.
Indeed, this model takes into account the physical phenomena that describe the behavior of the material in the formulation of the relationships between the stress $\sigma^y$, the strain $\varepsilon$, the strain rate $\mdot\varepsilon$ and the temperature $T$ expressed as power law, exponential law and hyperbolic sine.
This makes it easier to describe the softening phenomenon observed in the material due to increasing temperature.
The following equations describe the Arrhenius model:
\begin{equation}
\mdot\varepsilon =
\begin{cases}
A_1 \sigma^{y^{n_1}} \exp{\left(-\frac{Q}{RT}\right)} & \alpha\sigma^y < 0.8\\
A_2 \exp(\beta\sigma^y) \exp{\left(-\frac{Q}{RT}\right)} & \alpha\sigma^y > 1.2\\
A_3 \left[\sinh{\left(\alpha\sigma^y\right)}\right]^{n_2} \exp{\left(-\frac{Q}{RT}\right)} & \text{for all }\sigma^y
\end{cases}
\label{eq:ArDef}
\end{equation}
with:
\begin{equation}
Z = \mdot\varepsilon \exp{\left(\frac{Q}{RT}\right)}, \label{eq:ArZ}
\end{equation}
Where $Z$ is the Zenner-Hollomon parameter \cite{Zener-1944}, $\mdot\varepsilon$ is the strain rate ($\ps$), $Q$ is the apparent activation energy ($\text{J~mol}^{-1}$), $R$ is the universal gas constant ($8.314~\text{J~mol}^{-1} \text{K}^{-1}$), $T$ is the absolute temperature (K), $\sigma^y$ is the flow stress (MPa) for a given strain, strain rate and temperature and $A_1$, $A_2$, $A_3$, $n_1$, $n_2$, $\alpha$ and $\beta=\alpha~n_1$ are dependent of the material. The corresponding values are independent of the temperature and are obtained from the stress-strain curves at different strain rates and temperature by regression method. Combination of Equations (\ref{eq:ArDef}) and (\ref{eq:ArZ}) allow expressing the flow stress $\sigma^y$ as a function of the $Z$ parameter:
\begin{equation}
\sigma^y = \frac{1}{\alpha} \ln\left\{\left(\frac{Z}{A}\right)^{1/n} + \left[1 + \left(\frac{Z}{A}\right)^{2/n}\right]^{1/2}\right\}
\end{equation}

To obtain the constitutive equation, it is necessary to determine all the parameters $A$, $Q$ $\alpha$ and $n$ for a given material.
The strain has a significant non-linear influence on the behavior of the material by the mechanism of strain hardening and softening at high values of deformation.
A compensation in strain must therefore be taken into account in the Arrhenius model, which leads to the definition of the modified Arrhenius model for which the $A$, $Q$ $\alpha$ and $n$ parameters are expressed as a function of the strain $\varepsilon$ by means of polynomial functions of degree $m$ (varying from $1$ to $9$) of the form:
\begin{equation}
A(\varepsilon) = \exp{\left[\ln\!A_0 + \ln\!A_1\varepsilon + \ln\!A_2\varepsilon^2 + \ln\!A_3\varepsilon^3 + \cdots + \ln\!A_m\varepsilon^m\right]}
\label{eq:ArA}
\end{equation}
\begin{equation}
Q(\varepsilon) = Q_0 + Q_1\varepsilon + Q_2\varepsilon^2 + Q_3\varepsilon^3 + \cdots + Q_m\varepsilon^m
\label{eq:ArQ}
\end{equation}
\begin{equation}
\alpha(\varepsilon) = \alpha_0 + \alpha_1\varepsilon + \alpha_2\varepsilon^2 + \alpha_3\varepsilon^3 + \cdots + \alpha_m\varepsilon^m
\label{eq:Aralpha}
\end{equation}
\begin{equation}
n(\varepsilon) = n_0 + n_1\varepsilon + n_2\varepsilon^2 + n_3\varepsilon^3 + \cdots + n_m\varepsilon^m
\label{eq:Arn}
\end{equation}

The determination of the order $m$ of the polynomials defined by Equations (\ref{eq:ArA}-\ref{eq:Arn}) is defined according to the ability of the model to represent the non-linear dependence of the stress versus the strain and its generalization.
The values $\ln\!A_i$, $\alpha_i$, $n_i$, and $Q_i$ ($i=0,1,2,3,\cdots,m$) are the coefficients of the polynomials to determine using a regression method.
Setting $m=9$ gives the best results and the corresponding parameters are reported in Table \ref{tab:AR}.
\begin{table}[h!]
\centering
\caption{Parameters' values of the Arrhenius flow law for the P20 steel.}
\begin{tabular}{llll}
	\hline
	\multicolumn{1}{c}{$\alpha_i$}  & \multicolumn{1}{c}{$Q_i~(\times 10^{-6})$} & \multicolumn{1}{c}{$\ln\!A_i$} & \multicolumn{1}{c}{$n_i$}  \\ \hline
	$\alpha_0=0.0407$               & $Q_0=0.467$                                & $A_0=35.8092$                  & $n_0=4.8217$               \\
	$\alpha_1=-0.5167$              & $Q_1=-0.6517$                              & $A_1=-58.822$                  & $n_1=3.2814$               \\
	$\alpha_2=6.3912$               & $Q_2=7.6084$                               & $A_2=740.3303$                 & $n_2=71.5963$              \\
	$\alpha_3=-47.3364$             & $Q_3=-48.016$                              & $A_3=-5.0493\times 10^{3}$     & $n_3=-1.9562\times 10^{3}$ \\
	$\alpha_4=220.0014$             & $Q_4=66.795$                               & $A_4=1.1305\times 10^{4}$      & $n_4=1.4461\times 10^{4}$  \\
	$\alpha_5=-654.4553$            & $Q_5=468.8898$                             & $A_5=2.022\times 10^{4}$       & $n_5=-5.431\times 10^{4}$  \\
	$\alpha_6=1.2421\times 10^{3}$  & $Q_6=-2.3032\times 10^{3}$                 & $A_6=-1.5387\times 10^{5}$     & $n_6=1.1761\times 10^{5}$  \\
	$\alpha_7=-1.4523\times 10^{3}$ & $Q_7=4.3707\times 10^{3}$                  & $A_7=3.1798\times 10^{5}$      & $n_7=-1.4882\times 10^{5}$ \\
	$\alpha_8=952.0619$             & $Q_8=-3.9394\times 10^{3}$                 & $A_8=-2.9725\times 10^{5}$     & $n_8=1.0239\times 10^{5}$  \\
	$\alpha_9=-267.4994$            & $Q_9=1.397\times 10^{3}$                   & $A_9=1.0759\times 10^{5}$      & $n_9=-2.9621\times 10^{4}$ \\ \hline
\end{tabular}
\label{tab:AR}
\end{table}
This modified form of the Arrhenius behavior law allows for accurate and reliable prediction over a wide range of temperatures and strain rates.
Equation (\ref{eq:ArStress}) is finally used to compute the flow stress $\sigma^y$ from the strain $\varepsilon$, the strain rate $\mdot\varepsilon$ and the temperature $T$:
\begin{equation}
\sigma^y = \frac{1}{\alpha(\varepsilon)} \ln\left\{\left(\frac{\mdot\varepsilon \exp{\left(\frac{Q(\varepsilon)}{RT}\right)}}{A(\varepsilon)}\right)^{\frac{1}{n(\varepsilon)}} + \left[1 + \left(\frac{\mdot\varepsilon \exp{\left(\frac{Q(\varepsilon)}{RT}\right)}}{A(\varepsilon)}\right)^{\frac{2}{n(\varepsilon)}}\right]^{\frac{1}{2}}\right\}
\label{eq:ArStress}
\end{equation}

Figure \ref{fig:CompExp-AR-6} shows the comparison of the values predicted by the Arrhenius model and the experimental values.
The difference between the experimental and predicted values is small.
But for the strain rate $\mdot\varepsilon = 0.01~\ps$ and for the two low temperature values, the AR model is not able to predict the softening.
For the AR model, $\AARE=3.56~\%$ and $\RMSE=2.18~\MPa$.

\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-AR-6}
\caption{Comparison between the experimental and predicted flow stresses by AR model}
\label{fig:CompExp-AR-6}
\end{figure}

%----------------------------------------------------------------------------------
\subsection{The PTM model\label{sec:PTM}}
%----------------------------------------------------------------------------------

The PTM model \cite{TizeMha-2022} is a generalized formulation of the MZA model presented in Section \ref{sec:MZA}.
When establishing its formulation, the main shortcomings of the MZA model were taken into account, in order to make the PTM model flexible for any type of material studied because there is no need for a limited number of parameters as in the MZA model.
Its construction is based on the use of polynomial functions as in the AR model.
Thus, the intrinsic functions of the model, of polynomial type, allow adjusting the model according to the degree selected for each of the 4 constituent polynomials, which provides a good fit for each function.
The equation describing the PTM model is therefore given by :
\begin{equation}
\label{eq:PTM-model}
\sigma^y = \left(\sum_{i=0}^{q}{A_i\varepsilon^{p^i}}\right) \exp\left[\left(\sum_{j=0}^{r}{B_j\varepsilon^{p^j}}\right)\left(T-T_0\right) + \left(\sum_{k=0}^{s}\left(\sum_{l=0}^{t}{C_k^l\varepsilon^{p^l}} \right)\left(T-T_0\right)^k \right)\ln\left( \frac{\mdot\varepsilon}{\mdot{\varepsilon}_0}\right)\right]
\end{equation}
where $A_i, B_j, C_k^l$ are the parameters (Table \ref{tab:PTM}) of the model to be identified using the procedure proposed in  \cite{TizeMha-2022}.
Quantities $q$, $r$, $s$ and $t$ define the degree of the polynomials used to describe the behavior of the material.
The larger these quantities are, the more parameters need to be identified for the PTM model.
The determination of the parameters of this model are calculated thanks to the LMFIT Python library \cite{Newville-2016} and for more details about this model  we invite the reader to read our previous work \cite{TizeMha-2022}.
Thus, all the parameters of this model are calculated with $q=5$, $r=5$, $s=1$ and $t=5$.

Comparison of predicted values of the PTM model and experimental values is shown in Figure \ref{fig:CompExp-PTM-6}.
The PTM model is suitable to describe the flow behavior of medium carbon steel, but for the two strain rates $\mdot\varepsilon=0.01~\ps$ and $\mdot\varepsilon=0.1~\ps$ the prediction is not quite good.
The deviation between the predicted values and the experimental values for other strain rates is relatively good.
For the PTM model, $\AARE=4.79~\%$ and $\RMSE=4.59~\MPa$ and this shows an overall worse performance than the AR model.

\begin{table}[h!]
\centering
\caption{Parameters' values of the PTM flow law for the P20 steel.}
\begin{tabular}{llll}
	\hline
	\multicolumn{1}{c}{$A_i$}  & \multicolumn{1}{c}{$B_i$}   & \multicolumn{1}{c}{$C_0^i$} & \multicolumn{1}{c}{$C_1^i$}   \\ \hline
	$A_0=16.8529$              & $B_0=-3.5418\times 10^{-3}$ & $C_0^0=0.1608$              & $C_1^0=-1.9037\times 10^{-5}$ \\
	$A_1=340.6451$             & $B_1=-0.0132$               & $C_0^1=-0.6202$             & $C_1^1=2.67\times 10^{-3}$    \\
	$A_2=-1.9594\times 10^{3}$ & $B_2=-4.4888\times 10^{-3}$ & $C_0^2=4.9516$              & $C_1^2=-3.5788\times 10^{-3}$ \\
	$A_3=4.836\times 10^{3}$   & $B_3=0.2218$                & $C_0^3=-13.1694$            & $C_1^3=-0.0222$               \\
	$A_4=-5.5176\times 10^{3}$ & $B_4=-0.4988$               & $C_0^4=15.25$               & $C_1^4=0.0609$                \\
	$A_5=2.4058\times 10^{3}$  & $B_5=0.3211$                & $C_0^5=-6.587$              & $C_1^5=-0.0413$               \\ \hline
\end{tabular}
\label{tab:PTM}
\end{table}

\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-PTM-6}
\caption{Comparison between the experimental and predicted flow stresses by PTM model}
\label{fig:CompExp-PTM-6}
\end{figure}

%----------------------------------------------------------------------------------
\subsection{The Artificial Neural Network model\label{sec:ANNmodel}}
%----------------------------------------------------------------------------------

Because of their predictive capacity and their adaptability, Artificial Neural Networks (ANNs) are more and more widely used today in many scientific fields.
Their operation is based on a learning process during which the principle of minimizing the error between the model's output and the training data allows the adjustment of the model's parameters, as in any machine learning process.

There are generally two uses of neural networks: classification and regression.
The first is the ability to classify data into different groups and for example to distinguish between images of cats and dogs.
The second one corresponds to the capacity of universal approximator of these neural networks, which interests us in this application, and thus to the ability after learning to predict the values of the flow stress $\sigma^y$ according to the input data, as do the analytical models previously identified.
The main difference is that this approximation is not linked to a fixed mathematical formulation (JC, MZA, AR, HS, PTM model), but is only dependent on the data used for training, the number of layers, the number of neurons per layer and the activation functions associated with the neurons of the network.
A feed-forward ANN, as used in our application, contains an input layer, an output layer and a number of hidden layers ($2$ in our case).
Each layer of neurons is connected to the one before it and the one after it by weighted connections.
Thus, all the neurons of the $k$ layer are connected to all the neurons of the ($k-1$) layer as shown in Figure \ref{fig:ANN-2HL}.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\columnwidth]
{Figures/ANN-scheme-2HL}
\caption{Multi-layer Artificial Neural Network architecture}
\label{fig:ANN-2HL}
\end{figure}

Any hidden layer $k$, containing $n$ neurons, takes a weighted sum of the outputs $\overrightarrow{\hat{y}}$ of the immediately previous layer $(k-1)$, containing $m$ neurons, given by the following equation:
\begin{equation}
y_i\lay{k} = \sum_{j=1}^m w_{ij}\lay{k} \hat{y}_j^{(k-1)}+ b_i\lay{k},\label{eq:ANN1}
\end{equation}
where $y_i\lay{k}$ is the entry of the $i^{th}$ neuron of layer $k$, $\hat{y}_j\lay{k-1}$ is the output of the $j^{th}$ neuron of layer $(k-1)$, $w_{ij}\lay{k}$ is the associated weight parameter between the $i^{th}$ neuron of layer $k$ and the $j^{th}$ neuron of layer $(k-1)$ and $b_i\lay{k}$ is the associated bias of the $i^{th}$ neuron of layer $k$.
Those weights $w_{ij}$ and bias $b_i$, for each layer, are the training parameters of the ANN that we have to adjust during the training procedure described in Pantalé \eal \cite{Pantale-2021}.
For the proposed model, we have selected the Sigmoid activation function, so that, each neuron in the hidden layer $k$ provides an output value ${\hat{y}}$ from the input value $y$ of the same neuron defined by Equation (\ref{eq:ANN1}) according to the following equation:
\begin{equation}
\hat{y}=\frac{1}{1 + \e{-y}}\label{eq:ANN2}
\end{equation}
No activation function is used for the output neuron of the ANN.

After some tests of different types of network architecture and in accordance with previous works, a network structure with two hidden layers including $15$ neurons for the first hidden layer and 7 neurons for the second layer give the best compromise between prediction, learning time and model compactness.
From the global architecture point of view, the input layer is composed of three neurons ($\varepsilon^p$, $\mdot\varepsilon$, $T$) and the output layer is composed of a single neuron corresponding to the $\sigma^y$ flow stress.
This architecture leads to a global model with $180$ parameters to be identified ($60$ for the first layer, $112$ for the second layer and 8 for the output layer).

The Python program used for training the neural network was developed using the specialized Python library Tensorflow \cite{Abadi-2016}.
The Adaptive Moment Estimation (ADAM) optimizer \cite{Kingma-2015} was used for the training phase.
The training data are those from the tests presented in section \ref{sec:ComTestResults} and composed of $21~000$ quadruplets of ($\varepsilon^p$, $\mdot\varepsilon$, $T$, $\sigma^y$) values.
The learning is performed on the basis of $5~000$ epochs of the experimental data set.
Fourty minutes of training on a Dell XPS-13 7390 laptop running Ubuntu 22.04 LTS 64 bits with 16 GB of Ram and an Intel 4-core i7-10510U processor allow obtaining the converged parameters of the ANN model.
Figure \ref{fig:ANN-6-conv} show the evolution of the training error defined by the $\log_{10}$ of the $\RMSE$  during the training phase.
\begin{figure}[!ht]
\centering
\includegraphics[width=0.7\columnwidth]
{Figures/Conv-ANN-6}
\caption{Convergence of the ANN model during the training phase.}
\label{fig:ANN-6-conv}
\end{figure}
As we can see on this figure, after $5~000$ epoch, we can consider that we have reached a stationary state of the model learning and that it is useless to continue the learning phase.

Once the learning phase is over, the trained model can be used to predict the behavior of the medium carbon alloy as a function of the input data in the same way as is done with the analytical models.
One can either use the model directly by providing it with new input data, or retrieve the $180$ parameters identified during the training and inject them into a mathematical model based on Equations (\ref{eq:ANN1}) and (\ref{eq:ANN2}) which can be implemented in any language (\eg in FORTRAN for use on the Abaqus Explicit FEM code) as proposed in Pantalé \eal \cite{Pantale-2021}.
For reasons of compactness, the parameters of the ANN model and the complete procedure to compute the flow stress $\sigma^y$ from the input data are provided in the Appendix.

As for the analytical models presented previously, Figure \ref{fig:CompExp-3-15-7-1-sigmoid} shows a comparison between the flow stresses predicted by the ANN model and the data measured during the hot compression test.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-3-15-7-1-sigmoid}
\caption{Comparison between the experimental and predicted flow stresses by the 3-15-7-1-sigmoid ANN model}
\label{fig:CompExp-3-15-7-1-sigmoid}
\end{figure}
The correlation between the experimental data and the ANN prediction is very good over the entire range of data and the predicted data can track the hardening and softening regions of the hot deformed material well.
For the ANN model, $\AARE=0.62~\%$ and $\RMSE=0.38~\MPa$ which is excellent.
This model can be used to simulate the hot deformation of this type of alloy with much greater fidelity to the actual material behavior than the analytical models presented previously.

%----------------------------------------------------------------------------------
\subsection{Comparison of analytical and ANN models\label{sec:Comparison}}
%----------------------------------------------------------------------------------

A summary of the coefficients for evaluating the high-temperature flow stress prediction capability of the medium carbon alloy for all models presented in this work is reported in Table \ref{tab:Errors}.
\begin{table}[h!]
\centering
\caption{Accuracy coefficients for all the analyzed models.}
\begin{tabular}{lcccccc}
	\hline
	Coefficients        &   JC    &   MZA   &   HS   &   AR   &  PTM   &  ANN   \\ \hline
	$\AARE(\%)$         & $14.05$ & $21.20$ & $7.75$ & $3.56$ & $4.79$ & $0.62$ \\
	$\RMSE(\MPa)$ & $12.00$ & $19.57$ & $3.80$ & $2.18$ & $4.59$ & $0.38$ \\ \hline
\end{tabular}
\label{tab:Errors}
\end{table}
From this table, we can see that the ANN model has a much better predictive capacity than all the analytical models presented previously.The values of $\AARE$ and $\RMSE$ are globally 6 times lower than the best of the analytical models, \ie the Arrhenius model, quoted as reference in the context of hot forming of alloys \cite{Liang-2022}.

The ANN, Arrhenius and PTM models are the only models that take into account the softening with the deformation of medium carbon at low strain rate, unlike the other three models that only present an increase of the flow stress with the strain, whatever the strain rate and temperature, hence their poor performance in predicting the behavior of this material and more particularly at low strain rates.
The parameters reported in Table \ref{tab:Errors}, the correlations visible in Figures \ref{fig:CompExp-JC-6} to \ref{fig:CompExp-3-15-7-1-sigmoid} allow us to conclude here that the ANN model is the most efficient of all the models presented to describe the behavior of the medium carbon alloy for high temperature deformation applications.

%----------------------------------------------------------------------------------
\section{Interpolation and extrapolation capability of analytical and ANN models \label{sec:IntExt}}
%----------------------------------------------------------------------------------

In order to better compare the performances of the models presented previously (the 5 analytical models and the neural network based model), in this section, we present the ability of each of these models to interpolate and extrapolate the results as a function of the strain rate $\mdot\varepsilon$.
The identification of the models performed without the section is based on a set of experimental data corresponding to 6 strain rates, 5 temperatures and strains ranging from $0$ to $0.7$.
In order to test the learning capabilities and fidelity of these models, we propose here to perform the training of the models on only 5 strain rates by voluntarily omitting the strain rate $\mdot\varepsilon=1~\ps$ or $\mdot\varepsilon=5~\ps$.

Thus, one of the omitted strain rates ($\mdot\varepsilon=1~\ps$) is within the range of strain rates for model identification ($0.001~\ps\text{ to }5~\ps$), so we can test the ability of the models to interpolate the results from the other 5 strain rates and be able to quantify the deviation from the experimental values.
On the other hand, the omitted strain rate ($\mdot\varepsilon=5~\ps$) is outside the range of strain rates of identification of the models ($0.001~\ps\text{ to }2~\ps$), we will then test the ability of the models to extrapolate the results and also be able to quantify the deviation with the experimental values.

%----------------------------------------------------------------------------------
\subsection{Interpolation validation}
%----------------------------------------------------------------------------------

For the validation of the interpolation, the chosen omitted strain rate is $\mdot\varepsilon=1~\ps$ and those used for identification (or training for ANN) are the 5 other, \ie $\mdot\varepsilon=[0.001, 0.01, 0.1, 2, 5]~\ps$.
All models were re-identified from the same experimental data on a data set corresponding to 5 strain rates and 5 temperatures.

Figure \ref{fig:CompInt} shows a comparison, for strain rate $\mdot\varepsilon=1~\ps$, of the flow stresses $\sigma^y$ calculated by the models (as a line) and the experimental results (as points) for the 5 analytical models and the neural network.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompInt}
\caption{Comparison between the experimental (dot) and predicted flow stresses $\sigma^y$ (lines) for $\mdot\varepsilon=1~\ps$}
\label{fig:CompInt}
\end{figure}
Table \ref{tab:IntVal} shows the $\AARE$ and $\RMSE$ deviations between the different models and the experimental data calculated either for the 6 strain rates (lines referred as: all) or only on the strain rate $\mdot\varepsilon=1~\ps$.
\begin{table}[h!]
\centering{}
\caption{Accuracy coefficients of interpolation for all models with  $\mdot\varepsilon=1~\ps$}
\begin{tabular}{llcccccc}
	\hline
	Strain rate                      & Coefficients        &   JC    &   MZA   &   HS   &   AR   &  PTM   &  ANN   \\ \hline
	\mr{2}{id. $\mdot\varepsilon$}   & $\AARE(\%)$         & $13.79$ & $20.22$ & $8.57$ & $3.96$ & $5.10$ & $0.70$ \\
	                                 & $\RMSE(\MPa)$ & $11.83$ & $19.00$ & $3.97$ & $2.29$ & $4.73$ & $0.38$ \\ \hline
	\mr{2}{$\mdot\varepsilon=1~\ps$} & $\AARE(\%)$         & $14.87$ & $27.79$ & $3.74$ & $1.46$ & $3.16$ & $2.47$ \\
	                                 & $\RMSE(\MPa)$ & $12.02$ & $23.67$ & $3.10$ & $1.46$ & $2.65$ & $2.77$ \\ \hline
	\mr{2}{all $\mdot\varepsilon$}   & $\AARE(\%)$         & $14.42$ & $28.90$ & $7.55$ & $3.45$ & $4.90$ & $0.96$ \\
	                                 & $\RMSE(\MPa)$ & $11.86$ & $19.86$ & $3.84$ & $2.17$ & $4.45$ & $1.18$ \\ \hline
\end{tabular}
\label{tab:IntVal}
\end{table}

In Figure \ref{fig:CompInt} it can be seen that the first two models (JC and MZA), presented in this study do not have the ability to reproduce the behavior of the material for the strain rate $\mdot\varepsilon=1~\ps$.
Overall, the JC model gives better results than the MZA model for $\mdot\varepsilon=1~\ps$, which is reflected in Table \ref{tab:IntVal} by a lower value of $\AARE$ and $\RMSE$ for JC than for MZA.
Nevertheless, these values are higher than $10\%$ for the JC model and $20\%$ for the MZA model, which reflects the poor ability of these models to correctly model the behavior of this material.
This finding is in agreement with the previous findings of Sections \ref{sec:JC} and \ref{sec:MZA} which showed the inability of these models to take into account the softening of the behavior visible at low strain rates and low temperatures.

From a general appearance point of view, the other 4 models, HS, AR, PTM and ANN, give globally similar results with a higher fidelity for the AR and ANN models compared to the other two models.
Table \ref{tab:IntVal} shows the quantitative comparison on $\AARE$ and $\RMSE$ for three different cases: on the 5 identified strain rates (named id. $\mdot\varepsilon$), on only the strain rate $\mdot\varepsilon=1~\ps$ and on all 6 strain rates for these 6 models.
It appears from this Table that if the two models AR and ANN give equivalent (and excellent) results concerning the values of $\AARE$ and $\RMSE$ for the strain rate $\mdot\varepsilon=1~\ps$, the ANN model gives globally a better result on the whole of the strain rates with values of $\AARE=0.96\%$ and $\RMSE=1.18~\MPa$ respectively, that is to say values approximately 2 to 3 times lower for the ANN model compared to the AR model.

This shows the superior fidelity of the ANN model compared to the 5 analytical models presented in this study both in the interpolation capability of the model and in the overall behavior.

%----------------------------------------------------------------------------------
\subsection{Extrapolation validation}
%----------------------------------------------------------------------------------

For validation of the models' ability to extrapolate, the omitted strain rate chosen is $\mdot\varepsilon=5~\ps$ and those used for identification are $\mdot\varepsilon=[0.001, 0.01, 0.1, 1, 2]~\ps$.
The strain rate omitted in this analysis is therefore the highest value, which restricts the learning domain.

In this new identification configuration, Figure \ref{fig:CompExt} shows a comparison, for strain rate $\mdot\varepsilon=5~\ps$, of the flow stresses $\sigma^y$ computed by the models (as a line) and the experimental results (as points) for the 5 analytical models and the neural network.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExt}
\caption{Comparison between the experimental (dot) and predicted flow stresses $\sigma^y$ (lines) for $\mdot\varepsilon=5~\ps$}
\label{fig:CompExt}
\end{figure}
Table \ref{tab:ExtVal} shows the $\AARE$ and $\RMSE$ deviations between the different models and the experimental data calculated either for the 6 strain rates (lines referred as: all) or only on the strain rate $\mdot\varepsilon=5~\ps$.
\begin{table}[h!]
\centering{}
\caption{Accuracy coefficients of extrapolation for all models with  $\mdot\varepsilon=5~\ps$}
\begin{tabular}{llcccccc}
	\hline
	Strain rate                      & Coefficients        &   JC    &   MZA   &   HS   &   AR   &   PTM   &  ANN   \\ \hline
	\mr{2}{id. $\mdot\varepsilon$}   & $\AARE(\%)$         & $13.09$ & $19.09$ & $7.30$ & $4.03$ & $4.34$  & $0.61$ \\
	                                 & $\RMSE(\MPa)$ & $9.86$  & $16.26$ & $3.36$ & $2.32$ & $3.63$  & $0.32$ \\ \hline
	\mr{2}{$\mdot\varepsilon=5~\ps$} & $\AARE(\%)$         & $20.16$ & $24.27$ & $7.73$ & $3.53$ & $11.46$ & $3.87$ \\
	                                 & $\RMSE(\MPa)$ & $20.73$ & $25.95$ & $7.83$ & $4.02$ & $12.91$ & $5.84$ \\ \hline
	\mr{2}{all $\mdot\varepsilon$}   & $\AARE(\%)$         & $15.12$ & $25.87$ & $7.12$ & $3.86$ & $5.34$  & $1.09$ \\
	                                 & $\RMSE(\MPa)$ & $12.36$ & $18.24$ & $4.43$ & $2.68$ & $6.23$  & $2.40$ \\ \hline
\end{tabular}
\label{tab:ExtVal}
\end{table}

As presented in the previous section regarding the interpolation capability of the models, in Figure \ref{fig:CompExt} it can be seen that the first two models, JC and MZA, in this study again do not have the ability to correctly reproduce the material behavior for the strain rate $\mdot\varepsilon=5~\ps$.
The JC model performs better than the MZA model for $\mdot\varepsilon=5~\ps$, which is reflected in Table \ref{tab:ExtVal} by a lower value of $\AARE$ and $\RMSE$ for JC than for MZA.
Nevertheless, these values are too large for proper use.
It is again the inability of these models to take into account the softening of the behavior at low strain rates and low temperatures that is at the origin of these values.

The other 4 models, HS, AR, PTM and ANN, give better results with higher fidelity for the AR and ANN models compared to the other two models.
Table \ref{tab:ExtVal} shows the comparison over all strain rates and over $\mdot\varepsilon=5~\ps$ for these 6 models.
The HS model performs worse on extrapolation than the AR and ANN models, while the PTM model is relegated to the last position in this ranking with values of $\AARE$ and $\RMSE$ superior to $10\%$ for the strain rate $\mdot\varepsilon=5~\ps$.
The two models AR and ANN give the best results concerning the values of $\AARE$ and $\RMSE$ for the strain rate $\mdot\varepsilon=5~\ps$.
The AR model is this time a little bit better on the $\mdot\varepsilon=5~\ps$ strain rate compared to the ANN model, but the latter gives globally a better result on the whole strain rates with values of  $\AARE=1.09\%$ and $\RMSE=2.40~\MPa$ respectively.

We can therefore conclude from this part of the study that the AR model is the best performing of the 5 analytical models presented, which is in agreement with the fact that it is widely used for thermomechanical processing, but the ANN model approach has advantages over the AR model approach in that, overall, the ANN model is more faithful to the experimental data than the AR model for all strain rates.

%----------------------------------------------------------------------------------
\subsection{Proposed technique for better interpolation}
%----------------------------------------------------------------------------------
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/Performance}
\caption{Comparison between the experimental (dot) and predicted flow stresses $\sigma^y$ (lines) for $\mdot\varepsilon=1~\ps$}
\label{fig:Performance}
\end{figure}
In this part a technique is proposed to have a better interpolation of data. Indeed, as mentioned in the introduction, it is important to know how to reassure oneself of the effectiveness of the model that well describes the behavior of a material for a good use on new data. It is therefore a question of choosing models that have a good correlation with experience and proposing a formula for a good use of their parameters on unknown data. The AR, HS, PTM and ANN models are therefore chosen and it emerges that each model must be multiplied by a lambda factor whose formula is given as follows.
\begin{equation}
 \lambda = \left(\frac{\dot{\varepsilon}_{max} - \dot{\varepsilon}_{p} }{\dot{\varepsilon}_{p} - \dot{\varepsilon}_{min}}\right)^{\frac{\dot{\varepsilon}_p}{m}}
\end{equation}
Where $\dot{\varepsilon}_{p}$ is the unkown strain rate and $m\in[50-300]$
Applaying this formule the new figures are plotted in Figure \ref{fig:Performance}. From these results we can see clearly that the results are better that the initial ones. 
\section{Conclusions and future work\label{sec:Conclusion}}
%----------------------------------------------------------------------------------

Experimental tests were performed on a Gleeble thermomechanical simulator for a modified medium carbon carbon alloy to investigate the applicability and predictive accuracy of 5 analytical models and an Artificial Neural Network model over a range of strains ($0.0-0.7$), strain rates ($0.001~\ps-5~\ps$) and temperatures ($1050\celsius-1250\celsius$).
The analytical models selected for this study are the Johnson-Cook (JC) model \cite{Johnson-1983}, the Modified Zerilli-Armstrong (MZA) model \cite{Samantaray-2009}, the Hansel-Spittle (HS) model \cite{Hensel-1978}, the Arrhenius (AR) model \cite{Sellars-1966} and the PTM model  \cite{TizeMha-2022}.
Analysis of the data from the Gleeble trials, and comparison of the 6 models proposed in this study against the experiments, led to the following conclusions.

From an experimental point of view, it appears from the tests carried out that the flow stress $\sigma^y$ increases with the decrease of the temperature $T$ and the increase of the strain rate $\mdot\varepsilon$ due to the competitive appearance of the dynamic softening and work hardening mechanisms.
The dynamic recrystallization (DRX) phenomenon, introduced through the difference between the maximum and permanent strain, shows a partially complete microstructure evolution.
Thus, at high strain rates, it is difficult to visualize the DRX phenomenon on the flow curves due to the sensitivity of this phenomenon to the strain rate.
A study focused on an in-depth analysis of the microstructure of this steel alloy and its impact on mechanical properties is currently underway.

Five analytical models have been identified for this alloy, as well as an Artificial Neural Network based model.
Among the analytical models, the JC and MZA models proved to be inappropriate to reproduce the behavior of this material, while the HS, PTM and AR models showed their capabilities since they present an acceptable $\AARE$ error (from $3.5\%$ to $7.7\%$), with a superiority of the AR model ($3.56\%$) over the other two which confirms its use in thermomechanical processes.
The ANN model, is largely more accurate than the analytical models in predicting the flow stress $\sigma^y$ of the medium carbon with a value of $\AARE=0.62\%$.

To test the performance of each proposed model, a study was conducted to evaluate the interpolation and extrapolation capability of the developed models.
In the case of interpolation, the HS, PTM and AR models have a good correlation with the experiment, but the ANN model shows superiority with an error factor 5 times lower than the AR model.
For the extrapolation of the data, the HS, PTM and AR models have again a good correlation with the experiment, but the ANN model is again globally better performing.
These two parts of the study confirm what was demonstrated in the model identifications.

Identifying the parameters of an ANN model from experimental data requires more time than identifying the parameters of analytical models (about 40 minutes on a working laptop), but as shown by Pantalé \eal \cite{Pantale-2021}, implementing an ANN model in a computational code is straightforward.
Therefore, we recommend using the ANN model when conducting a study involving this alloy under these experimental conditions.

As part of the perspectives of this work, we will develop another ANN model to characterize this time the microstructure in terms of DRX and formation of phases such as ferrite, bainite and martensite for example.
This ANN model can then be compared to the most commonly used classical models such as the JMAK model.
Then, these ANN models will be implemented in FEM software such as Abaqus or Forge in order to validate them.

\bibliographystyle{elsarticle-num}
\bibliography{Bibliography}

%----------------------------------------------------------------------------------
\section*{Appendix\label{sec:Appendix}}
%----------------------------------------------------------------------------------

In order to complete this paper, we report here after the computing process and the $180$ coefficients of the Artificial Neural Network ANN-3-15-7-1-sigmoid model used in Section \ref{sec:ANNmodel}.
In order to use this model we describe here after the procedure to compute the flow stress $\sigma^y$ from the input variables $\varepsilon^p$, $\mdot\varepsilon$ and $T$.
This process can be decomposed in 3 phases:
\begin{itemize}
\item We first have to normalize the input values of the ANN $x_i$ within the range $[0,1]$ to avoid ill-conditioned system as presented by many other authors in the literature \cite{Lin-2008-ANN, Lu-2011-ANN}.
Therefore, the three components of the input vector $\overrightarrow{x}$ are obtained from the plastic strain $\varepsilon^p$, the plastic strain rate $\mdot{\varepsilon}^p$ and the temperature $T$ using the following expressions:
\begin{equation}
\overrightarrow{x}=
\begin{cases}
x_1 = \frac{\varepsilon^p - [\varepsilon^p]_{min}}{[\varepsilon^p]_{max} - [\varepsilon^p]_{min}}\\
x_2 = \frac{\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})-[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{min}}{[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{max}-[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{min}}\label{eq:CR1}\\
x_3 = \frac{T-[T]_{min}}{[T]_{max}-[T]_{min}}
\end{cases}
\end{equation}
where $[~]_{min}$ and $[~]_{max}$  are the boundaries of the range of the corresponding field: $\varepsilon^p\!\in\!\left[0.0,0.7\right]$, $\mdot{\varepsilon}\!\in\!\left[0.001~\ps,5.0~\ps\right]$, $T\!\in\!\left[1050~\celsius,1250~\celsius\right]$ and $\sigma\!\in\!\left[1.311~\MPa,153.739~\MPa\right]$.
The reference strain rate is $\mdot{\varepsilon_0} = 0.001~\ps$.
\item Then we compute the output $s$ of the ANN from the input vector $\overrightarrow{x}$ using the three following equations.
\begin{equation}
\overrightarrow{y}_1 = \left[1 + \exp{\left(- \w_1 \dotp \overrightarrow{x}- \overrightarrow{b}_1\right)}\right]^{-1}
\end{equation}
\begin{equation}
\overrightarrow{y}_2 = \left[1 + \exp{\left(- \w_2 \dotp \overrightarrow{y}_1- \overrightarrow{b}_2\right)}\right]^{-1}
\end{equation}
\begin{equation}
s = \overrightarrow{w}^T \dotp \overrightarrow{y}_2 + b
\end{equation}
\item Finally, the flow stress $\sigma^y$ can be obtained from the output $s$ of the ANN using the following equation:
\begin{equation}
\sigma^y =  \left([\sigma]_{max}-[\sigma]_{min}\right)s + [\sigma]_{min} \label{eq:CR2}
\end{equation}
\end{itemize}

Conforming to the computing process proposed by Equations (\ref{eq:CR1}-\ref{eq:CR2}), we report hereafter the $180$ coefficients of the ANN-3-15-7-1-sigmoid model used in Section \ref{sec:ANNmodel}.
The weight matrix for the first hidden layer $\w_1$ is a $15\times3$ matrix:
\begin{equation*}
\w_1 = \left[
\begin{array}{rrr}
2.2206 & -3.7555 & -6.7246\\
-4.8598 & 5.7431 & -5.8538\\
2.3099 & 3.3325 & -5.1795\\
2.0475 & 0.8006 & -1.4259\\
8.8358 & -6.0362 & 0.8226\\
-1.2613 & -0.9274 & -2.3725\\
-0.3561 & 6.5032 & -10.7573\\
-11.7226 & -2.0455 & 1.2248\\
3.1066 & 26.5580 & 18.6540\\
-0.5150 & -5.6922 & 1.0104\\
-6.4755 & 8.4888 & -2.4459\\
-1.8791 & -0.5380 & 2.2295\\
-6.0206 & 1.2776 & 0.2169\\
0.2619 & -4.7974 & -1.1282\\
-27.1456 & -0.5327 & -0.5303\\
\end{array}\right]
\end{equation*}
The biases of the first hidden layer $\overrightarrow{b_1}$ is a $15$ components vector:
\begin{equation*}
\overrightarrow{b}_1 = \left[
\begin{array}{r}
5.9164\\
-1.9074\\
-2.6837\\
-1.0954\\
-0.9509\\
3.1682\\
-4.3964\\
0.6343\\
-5.0676\\
2.0228\\
1.1267\\
-0.9671\\
-0.6263\\
2.3110\\
-0.3037\\
\end{array}\right]
\end{equation*}
The weight matrix for the second hidden layer $\w_2$ is a $7\times15$ matrix:
\begin{equation*}
\w_2^T = \left[
\begin{array}{rrrrrrr}
-0.5783 & 1.2724 & 0.5747 & 0.6449 & -4.2203 & -0.2380 & 0.2591\\
-0.4852 & 5.2807 & 0.8888 & -8.2324 & -2.0075 & -0.6474 & -1.0787\\
4.4499 & -0.0137 & 0.1657 & 0.3198 & 4.9765 & -1.2503 & 0.8219\\
1.7571 & 0.7730 & 0.0208 & -1.3316 & -0.8945 & -0.7284 & -0.1831\\
-1.0866 & 0.1330 & -0.8615 & -0.1283 & 0.2218 & -0.1772 & -2.7458\\
-0.3925 & 1.3994 & 0.0630 & -1.8397 & -1.1047 & -1.9839 & 0.5767\\
-0.2121 & 0.9977 & 1.2028 & -9.6525 & 0.5520 & 0.1062 & -0.0409\\
-1.1518 & -1.6402 & -4.1501 & -1.0759 & 0.4749 & -2.8350 & 0.9225\\
-1.1453 & -0.2173 & -0.1382 & 0.8264 & -0.5125 & 0.1882 & -0.8654\\
-4.3301 & -0.3711 & -7.4305 & 3.5926 & -9.6217 & -1.2375 & 1.6171\\
2.3907 & -1.0085 & -0.8828 & -1.1891 & 0.9947 & 1.1178 & -1.0953\\
-1.5955 & 1.6313 & 0.4916 & 0.1906 & -1.9216 & -1.4140 & 1.3827\\
-1.6985 & 1.4277 & -3.4462 & -8.3777 & -1.2132 & -1.2158 & 2.8512\\
-1.9954 & -2.0159 & -8.3455 & -0.5205 & 0.2942 & -1.3337 & 0.2026\\
4.3040 & -0.7164 & -1.0859 & 3.4294 & -23.8003 & 12.5859 & 7.3721\\
\end{array}\right]
\end{equation*}
The biases of the second hidden layer $\overrightarrow{b_2}$ is a 7 components vector:
\begin{equation*}
\overrightarrow{b}_2 = \left[
\begin{array}{r}
0.7534\\
0.9473\\
0.6055\\
-0.7793\\
-1.0305\\
-1.5779\\
-0.1471\\
\end{array}\right]
\end{equation*}
The weight vector for the output layer $\overrightarrow{w}$ is a 7 components vector:
\begin{equation*}
\overrightarrow{w} = \left[
\begin{array}{r}
0.1920\\
0.3406\\
0.3839\\
-0.2880\\
1.2047\\
-1.4126\\
-0.2215\\
\end{array}\right]
\end{equation*}
The bias of the output layer $b$ is a scalar:
\begin{equation*}
b = 0.1178
\end{equation*}

\end{document}